---
title: "PML Project - Barbell Lifts"
author: "Anton Kiselev"
date: "21 Sep 2014"
output: html_document
---

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 

In this project we analyse the data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. The goal of the project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. For any purpose please cite them. 

Load and overview the data:
```{r}
tr <- read.csv("pml-training.csv")
te <- read.csv("pml-testing.csv")
summary(tr)
```

Clean data, select non-empty predictors only:
```{r}
predictors <- c("roll_belt", "pitch_belt", "yaw_belt", "total_accel_belt", "gyros_belt_x", "gyros_belt_y", "gyros_belt_z", "accel_belt_x", "accel_belt_y", "accel_belt_z", "magnet_belt_x", "magnet_belt_y", "magnet_belt_z", "roll_arm", "pitch_arm", "yaw_arm", "total_accel_arm", "gyros_arm_x", "gyros_arm_y", "gyros_arm_z", "accel_arm_x", "accel_arm_y", "accel_arm_z", "magnet_arm_x", "magnet_arm_y", "magnet_arm_z", "roll_dumbbell", "pitch_dumbbell", "yaw_dumbbell", "total_accel_dumbbell", "gyros_dumbbell_x", "gyros_dumbbell_y", "gyros_dumbbell_z", "accel_dumbbell_x", "accel_dumbbell_y", "accel_dumbbell_z", "magnet_dumbbell_x", "magnet_dumbbell_y", "magnet_dumbbell_z", "roll_forearm", "pitch_forearm", "yaw_forearm", "total_accel_forearm", "gyros_forearm_x", "gyros_forearm_y", "gyros_forearm_z", "accel_forearm_x", "accel_forearm_y", "accel_forearm_z", "magnet_forearm_x", "magnet_forearm_y", "magnet_forearm_z")
tr <- tr[,c(predictors,"classe")]
te <- te[,predictors]
```

Train random forest with cross-validation:
```{r, cache=TRUE}
set.seed(54321)
library(caret)
library(doMC)
registerDoMC(cores = 2)
ctrl <- trainControl(method = "cv", number = 5)
fit <- train(classe ~ ., data=tr, method = "rf", trControl = ctrl)
```

Result (expected error is 0.44%):
```{r}
fit$finalModel
plot(fit)
```

Making predictions and creating files:
```{r}
answers = predict(fit, newdata=te)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```
